Writing this thesis has been a long and challenging journey, one that I could not have completed without the help, encouragement, and inspiration of many people. I am deeply grateful to all those who supported me along the way.\\

I would first like to thank my parents, my best friend and father Ahmad, who sat me in front of a computer before I could walk. That simple gesture lit the spark that would later become a lifelong passion. And my mother Patricia, a model of perseverance and resilience, whose unwavering love and strength have carried me through more moments than I can count. This achievement is as much theirs as it is mine. To the rest of my family, who have always believed in me and encouraged me to pursue my dreams, thank you for your constant support and for always being there.\\

I am deeply grateful to Professor E. Solana, whose guidance and insight shaped this thesis and left a lasting mark on my academic journey. It was through one of your courses that I first discovered my passion for this field. Your thoughtful teaching and mentorship not only opened the door but gave me the confidence to walk through it. Thank you for your support, patience, and for always encouraging me to think deeper.\\

To my grandfathers Jose and Salah, both very different, both deeply inspirational. One, a stonemason, whose legacy of craftsmanship and hard work continues to inspire me. The other, an Arabic teacher, lover of literature and philosophy, who nurtured my love of learning and taught me the value of precision and clarity. I wish I could have spent more time with them both, but I carry their lessons with me every day.\\

There are so many others whose presence made a difference, even in small ways. I wish I could name every single one of you here. Please know that your support has not gone unnoticed, and I am sincerely thankful.\\